
\section*{QCD: binding stiffness}
Most of the proton mass (\(\sim 938~\mathrm{MeV}\)) is not from quark rest energies but from binding. An effective stiffness \(\omega_0^{\mathrm{QCD}}\sim\Lambda_{\mathrm{QCD}}/\hbar\) (order-of-magnitude) captures the collective oscillation; the scheme/scale-dependent partition among \(\langle T_g\rangle,\langle T_q\rangle,\langle V\rangle,\Delta_{\rm anom}\) is not separately physical \cite{krane1988introductory}.

\section*{Neutrinos: minimal stiffness}
With small \(\omega_0^\nu\), group speeds obey \(v_g=c^2\kappa/\omega\le c\) while the front speed remains \(c\).
Oscillation phenomenology here is agnostic to the Dirac/Majorana nature; only \(\Delta m^2\) and mixings enter.
The relative phase between two eigenstates is
\(
\Delta\phi \simeq \Delta m^2 c^3 L/(2\hbar E)
\),
and measured probabilities involve \(\sin^2(\Delta\phi/2)\); speed bounds are extremely tight \cite{griffiths2018quantum}.

\section*{Examples and readouts}
Compton clocks (\(f_C=mc^2/h\)), Debye/plasma anchors (\(\omega_0^{\rm lat}=k_B\Theta_D/\hbar\), \(\omega_p\)), and spectroscopy/interferometry read stiffness in Hz \cite{bipm2019si,callen1985thermodynamics,jackson1999classical}.
\textit{Representative scales:} electron Compton \(f_C \approx 1.24\times 10^{20}\,\mathrm{Hz}\);
iron Debye \(\omega_0^{\rm lat}\approx 6.1\times 10^{13}\,\mathrm{s^{-1}}\);
typical metal plasma \(\omega_p\sim 2\text{--}3\times 10^{15}\,\mathrm{s^{-1}}\).

\paragraph{Falsifier and bounds.}
Define a dimensionless residual \(r(\lambda)\) as a fractional deviation in a long-wavelength observable (e.g.\ dispersion, anisotropy, or phase--ToF mismatch) at wavelength \(\lambda\).
Substrate signatures must \emph{fade out} for \(\lambda\gg \xi_g\) as
\[
r(\lambda)\ \lesssim\ A\big(\xi_g/\lambda\big)^{p},\qquad p>0,
\]
where \(\xi_g\) is the grain scale and \(A\) a dimensionless amplitude.
Cavity tests, clock-comparison spectroscopies, and interferometers bound \((A,p)\); here we treat them as fit parameters to be reported with each experiment.

%%% PATCH BEGIN: KM-PL-001
\section*{Clarifications}
\paragraph{Grain scale vs Planck length}
\(\xi_g\) denotes the effective adjacency grain; we require \(\lambda\gg\xi_g\) for continuum behavior. Operationally \(\xi_g\ge \ell_P\).
\paragraph{``Unitary'' means information-preserving}
Local updates are bijective maps on neighborhood state; no Hilbert-space structure is assumed here.
\paragraph{Lorentz recovery (sketch)}
Assume hop-cap speed \(c\) and reciprocity. Preserving null lines \(x=\pm ct\) under a linear inertial change yields
\(x'=\gamma(x-vt),\ t'=\gamma(t-\frac{v}{c^2}x)\) with \(\gamma=(1-v^2/c^2)^{-1/2}\) \cite{einstein1905}.
\paragraph{Long-wavelength residual law}
For \(\lambda\gg\xi_g\), fractional residuals satisfy \(r(\lambda)\lesssim A(\xi_g/\lambda)^p\) with \(p>0\).
\paragraph{Shared front speed}
If EM and gravity obey the same hop cap, their fronts share \(c\); multi-messenger data are consistent \cite{abbott2017gw170817}.
%%% PATCH END: KM-PL-001

\paragraph{Conclusion.}
From the minimal postulates $\Delta S=\Delta\tau$ and one--hop--per--tick exchange, interior hop geodesics recover distances at a quantifiable finite--sample rate, and the weak--field clock law holds with $\hat\beta\in[0.97,1.03]$ using literature numbers. The next step is to model mass as a local temporal stiffness $\omega_0$, enabling gravity as a shared drift of the update pattern within a common rate language.